---
title: "How to use Ecological Metadata Language (EML)"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

## Test using the EML package

### First create an attribute metadatra

Create a table with the column names (attributeName) of a csv file (or similar) and fill in some other columns, the following table is an example

```{r, echo = TRUE}
library(EML)

attributes <-tibble::tribble(
~attributeName, ~attributeDefinition,                                                 ~formatString, ~definition,        ~unit,   ~numberType,
  "run.num",    "which run number (=block). Range: 1 - 6. (integer)",                 NA,            "which run number", NA,       NA,
  "year",       "year, 2012",                                                         "YYYY",        NA,                 NA,       NA,
  "day",        "Julian day. Range: 170 - 209.",                                      "DDD",         NA,                 NA,       NA,
  "hour.min",   "hour and minute of observation. Range 1 - 2400 (integer)",           "hhmm",        NA,                 NA,       NA,
  "i.flag",     "is variable Real, Interpolated or Bad (character/factor)",           NA,            NA,                 NA,       NA,
  "variable",   "what variable being measured in what treatment (character/factor).", NA,            NA,                 NA,       NA,
  "value.i",    "value of measured variable for run.num on year/day/hour.min.",       NA,            NA,                 NA,       NA,
  "length",    "length of the species in meters (dummy example of numeric data)",     NA,            NA,                 "meter",  "real")
```



```{r}
knitr::kable(attributes)
```

### then generate a key for factors

```{r, echo = T}
i.flag <- c(R = "real",
            I = "interpolated",
            B = "bad")
variable <- c(
    control  = "no prey added",
    low      = "0.125 mg prey added ml-1 d-1",
    med.low  = "0,25 mg prey added ml-1 d-1",
    med.high = "0.5 mg prey added ml-1 d-1",
    high     = "1.0 mg prey added ml-1 d-1",
    air.temp = "air temperature measured just above all plants (1 thermocouple)",
    water.temp = "water temperature measured within each pitcher",
    par       = "photosynthetic active radiation (PAR) measured just above all plants (1 sensor)"
)

value.i <- c(
    control  = "% dissolved oxygen",
    low      = "% dissolved oxygen",
    med.low  = "% dissolved oxygen",
    med.high = "% dissolved oxygen",
    high     = "% dissolved oxygen",
    air.temp = "degrees C",
    water.temp = "degrees C",
    par      = "micromoles m-1 s-1"
)

## Write these into the data.frame format
factors <- rbind(
    data.frame(
        attributeName = "i.flag",
        code = names(i.flag),
        definition = unname(i.flag)
    ),
    data.frame(
        attributeName = "variable",
        code = names(variable),
        definition = unname(variable)
    ),
    data.frame(
        attributeName = "value.i",
        code = names(value.i),
        definition = unname(value.i)
    )
)
```

which generates the following table

```{r}
knitr::kable(factors)
```

### we then generate an attributeList

```{r, echo=TRUE}
attributeList <- set_attributes(attributes, 
                                factors, 
                                col_classes = c("character", "Date", "Date", "Date", "factor", "factor", "factor", "numeric"))
```

### We link this to the csv file

```{r, echo = T}
physical <- set_physical("hf205-01-TPexp1.csv")

```

### And Assemble the dataTable

```{r, echo = T}
dataTable <- list(
                 entityName = "hf205-01-TPexp1.csv",
                 entityDescription = "tipping point experiment 1",
                 physical = physical,
                 attributeList = attributeList)
```

## Want to learn more about EML?

* go to this [link](https://eml.ecoinformatics.org/)

## External resources

* Good [reproduction list](https://www.youtube.com/watch?v=gW_-XTwJ1OA&list=PLi1PZkcSXdAJdmXchf9jvWMPZzsMoDVmA) on clean data 